---
title: Interview Presentation
subtitle: SRA (Genetic Epidemiology and Medical Statistics)
author: "[Paul Smith](mailto:psmith@posteo.net)"
date: 2024-12-12
date-format: "MMM D, YYYY"
format:
  revealjs:
    theme: serif
    fontsize: 2em
    df-print: paged # pages html tables: https://quarto.org/docs/computations/r.html#data-frames
execute:
  echo: false
logo: https://www.bristol.ac.uk/styles/t4/responsive/uob-logo.svg
---

# A Recent Project

```{r}
#| message: false
#| warning: false

here::i_am("index.qmd")

# loading packages
library(tidyverse)
library(patchwork) # for grid.arrange
library(lme4)
library(lmerTest)
library(rlang)
#library(flexmix) # for mixture models
library(gamlss) # for gamma multilevel modelling
```

```{r}

# loading data

load(here::here("data", "data.rda"))

# pulling out only the key variables for this model

data_craft <- tibble(data_craft) %>% 
    select(lengthstay, Log2lengthstay, Centre_ERAS_status, 
           donor_age, donor_type, CIT_mins, age, 
           who_performance_status, pre_emptivetx, 
           redcap_data_access_group)
```


## Background

- A retrospective cohort study investigating enhanced recover after surgery (ERAS) on (log) patient length of stay after kidney transplantation.
- ERAS is implemented at a centre-level.
- Eighteen hospitals (approx half have ERAS) and approx 200 patients.
- Working alongside a surgeon: the initial thought was to do a mixed model where centre is a random effect, and ERAS is a fixed binary effect

::: {.notes}
- ERAS: ERAS is a way of managing patients after a transplant, to try and get
  them discharged and back to normal activity as quickly as possible
- Structure of the data is:
    - Centre level data on whether a formal ERAS programme is in place. Along with some details of what this ERAS programme entails.
    - Patient level data on donor/recipient/operative details and outcomes.
- Confounding:
  - between the ERAS and centre effects,
  - also a human factor in all of this, with respect to how/ when ERAS is
    implemented in a centre.
- Blood and Transplant Research Units (BTRU) project: Organ donation and transplantation stream
:::

## Initial thoughts

A multilevel model for the log of length of stay.

::: {style="font-size: 50%;"}
```{r}
#| label: full-cohort-model

## This is a multilevel/mixed/hierarchical model for log2 length of stay. it has
# random intercepts based on redcap_data_access_group. Centre_ERAS_status is constant
# within each centre (redcap_data_access_group), and is therefore analysed at the 
# centre level, which deals with the unit of analysis issues that I was talking
# about previously
# merging the factors for cardiff and royal_free
data_craft$centre <- forcats::fct_collapse(data_craft$redcap_data_access_group, 
                        uhw_cardiff = c("uhw_cardiff", "uhw_cardiff_NoERAS"),
                        royal_free_london = c("royal_free_london_ERAS", 
                                                  "royal_free_london_NoERAS"))
```

```{r}
#| echo: true
model_main <- lmer(Log2lengthstay ~ Centre_ERAS_status + donor_age +
                     donor_type*CIT_mins + age + who_performance_status + 
                     pre_emptivetx + (1|centre), data = data_craft)
summary(model_main)
```
:::

::: {.notes}
- Linear Mixed-Effects Model output
- My job is to take a step back and discuss with them the limitations of the models and the data, and any confounding or reverse causation that might be occurring

:::

## Exploratory analysis

```{r}
# summary
#skimr::skim(data_craft)
# number of observations per centre
data_craft %>%
  group_by(redcap_data_access_group) %>%
  summarise("Number of Patients" = n())
```

::: {.notes}
With small group sizes, (e.g. fewer than 5-10 patients), there might be convergence issues when fitting the group-level effects. Additionally, the  variance of the group-level intercepts depends on the amount of data within each group.

- Q: Is there any way of aggregating the centres?
- How come both UHW Cardiff and Royal Free are split by ERAS and no ERAS, when ERAS is supposed to be done at a centre level?
:::

## Exploratory analysis: length of stay {.smaller}

:::: {.columns}

::: {.column width="40%"}

```{r}
#| layout-nrow: 2

# hist / density of length of stay
ggplot(data_craft, aes(x = lengthstay)) +
  geom_histogram(aes(y = after_stat(density)), bins = 10, 
                 fill = "#aa3377", alpha = 0.5) +
  geom_density(bw = 0.25) +
  theme_minimal() +
  labs(x = "LOS")

ggplot(data_craft, aes(x = Log2lengthstay)) +
  geom_histogram(aes(y = after_stat(density)), bins = 15, 
                 fill = "#aa3377", alpha = 0.5) +
  geom_density(bw = 0.15) +
  theme_minimal() +
  labs(x = "LOS (log-transformed)")
```

:::

::: {.column width="60%"}

::: {.fragment .fade-in}

Issues:

::: {.incremental}
- Clearly not Gaussian
  - Consider: generalised linear model with a gamma distribution
- Follow up was 30 days: *"Because we didn't want to lose these patients from the analysis, they are coded as having the 'maximum' length of stay for our follow up (i.e. 30 days)."*
  - Consider: Survival analysis *e.g.* Kaplan-Meier, Cox model, Frailty model
:::

:::

:::

::: {.fragment .fade-in}

```{r}
#| echo: true
#| eval: false
lmer(Log2lengthstay ~ Centre_ERAS_status + donor_age +
     donor_type*CIT_mins + age + who_performance_status + 
     pre_emptivetx + (1|centre), data = data_craft)
```

:::
:::

::: {.notes}
Clearly not Gaussian:

- GLM with a Gamma distribution: choosing a log-link function.
  - This means the (exponential of the) coefficients represent multiplicative effects on the length of stay.

**Steps:**

I think it should be simplified as much as possible, with the aim to get a minimum trustworthy model – this will reduce the chances of overfitting and make the findings more generalisable.

1. Can you plot a KM split by ERAS? See if a unadjusted model shows any difference in the survival curves. You will need to caveat that this is unadjusted for all other factors and that the ERAS effect could be moping up the centre effect also, but it is a good first ‘look’ at the data.
2. A Cox with frailty (centre) term and ERAS as the only fixed effect: as ERAS is so confounded with the centre, we don’t want to be interpreting ‘ERAS effect’ without the presence of centre.
3. Include one or two risk factors that you think (clinically) are the most important.
:::

# What will I bring to the job? 

## Skills 
::: {.incremental}
- Research: *masters, PhD and at NHSBT*
- Programming: *R, SAS, Python, and Stata*
- Teaching: *at Leeds University and a secondary school*
- Collaboration and working in interdisciplinary teams: *through the BTRU partnership*
- Passionate! 
  - Want to make a difference through my work and love working as a medical statistician: varied, interesting and impactful!

:::

::: {.notes}
**Research**

- 1st Class w Honors MMath at Bristol
- PhD from Leeds University: 
  - Forecasting large-scale complex simulators: classification, dimensional reduction, stochastic processes
- Research within NHSBT: worked in NIHR Blood and Transplant Research Units (BTRU) partnership with clinicians at Cambridge and Newcastle university

**Programming**

- Strong R skills, including using RAP to create reproducible, accurate analysis (e.g. by implementing Quarto)
- Use of version control and collaboration via git
- Also know: Python and SAS

**Teaching**

- Small group tutorial teaching at Leeds University whilst doing my PhD
- Two years of experience teaching at a secondary school and sixth form

**Collaboration**

:::


